---
title: "Data Cleaning"
author: "AnnaClaire Marley"
date: "6/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**This rmarkdown includes cleaning and pre-processing of raw data to use in our solar suitability model (model.Rmd)**

Load packages
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(raster)
library(maps)
library(sf)
```

```{r}
# Create a Root to Extract Files 
if(Sys.info()["sysname"] == "Windows"){ #If operating system is "Windows" use G: Drive
  team_path <- "G:/"
} else { # If it is not (i.e. Mac), use the Google Drive File Stream
  team_path <- "/Volumes/GoogleDrive/"
}

#Create the Full File Path
path <- paste0(team_path, "Shared drives/ESM232/Data/in")

```

**CRS**
```{r}
# Assign projection: NAD83
NAD83 <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83
+units=m +no_defs +ellps=GRS80 +towgs84=0,0,0 "
```

**Path to functions for cleaning data**
```{r}
clean_path <- "./Functions/data_cleaning/"
```

**Make California shapefile**

```{r}
## State border##
# grab US states from maps package
california <- st_as_sf(maps::map("state", plot = FALSE, fill = TRUE)) %>% 
  filter(ID == "california")

# make the polygon the same CRS as rasters
california_project <- st_transform(california, NAD83)

## counties ##
ca_counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE)) %>% 
  filter(str_detect(ID, 'california'))

# make the polygon the same CRS as rasters
county_project <- st_transform(ca_counties, NAD83)

```

## Raster Files cleaning -- Solar Energy Output Model

Data available for download from Cal-Adapt raster download: http://v1.cal-adapt.org/data/download/

Source Functions
```{r}
source(paste0(clean_path, "process_rasters.R")) # for raster processing
source(paste0(clean_path, "ca_crop_mask.R")) # for cropping, masking, reprojecting rasters
```

#### Net surface radiation (W/m^2)

* Model: CCSM3
* Emission scenario: A2
* Year: 2050

```{r}
# net average surface radiation 
solar_stack <- process_rasters(folder_path = "/solar_2020")

# crop and mask to ca
solar_stack_mask <- ca_crop_mask(solar_stack, NAD83, california_project)

# plot
plot(solar_stack_mask)

# save file
writeRaster(solar_stack_mask, 
            filename="./generated_files/solar_stack_mask_20.tif",
            options="INTERLEAVE=BAND", overwrite=TRUE)

```

#### Temperature (C)

* Model: CCSM3
* Emission scenario: A2

##### Average Daily 

```{r}
#generate a list of input rasters 
temp_stack <- process_rasters(folder_path = "/tav_2020")

# crop and mask to ca
temp_stack_mask <- ca_crop_mask(temp_stack, NAD83, california_project)

# plot
plot(temp_stack_mask)

# save file
writeRaster(temp_stack_mask, 
            filename="./generated_files/temp_stack_mask_20.tif",
            options="INTERLEAVE=BAND", overwrite=TRUE)
```

##### Average Maximum 

```{r}
# max temp
tmax_stack <- process_rasters(folder_path = "/tmax_2020")

# crop and mask to ca
tmax_stack_mask <- ca_crop_mask(tmax_stack, NAD83, california_project)

# plot
plot(tmax_stack_mask)

# save file
writeRaster(tmax_stack_mask, 
            filename="./generated_files/tmax_stack_mask_20.tif",
            options="INTERLEAVE=BAND", overwrite=TRUE)
```

#### Average Minimum 

```{r}
# max temp
tmin_stack <- process_rasters(folder_path = "/tmin_2020")

# crop and mask to ca
tmin_stack_mask <- ca_crop_mask(tmin_stack, NAD83, california_project)

# plot
plot(tmin_stack_mask)

# save file
writeRaster(tmin_stack_mask, 
            filename="./generated_files/tmin_stack_mask_20.tif",
            options="INTERLEAVE=BAND", overwrite=TRUE)
```


## Cleaning -- Land suitability Model


# cleaning -- Economic model




