---
title: "Model Example"
author: "AnnaClaire Marley"
date: "6/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Rmarkdown runs through the solar suitability model with sample data cleaned and processed in the data_cleaning.Rmd

Load libraries
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(raster)
library(maps)
library(sf)
library(paletteer)
library(stringr)
library(RColorBrewer)
```

**Path to functions for the model**
```{r}
model_path <- "./Functions/model/"
```

# CA counties

```{r}
## counties ##
ca_counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE)) %>% 
  filter(str_detect(ID, 'california'))

# make the polygon the same CRS as rasters
county_project <- st_transform(ca_counties, NAD83)
```

## Projected Solar Panel Energy Output Model

```{r}
# source function
source(paste0(model_path, "energy_output.R"))
```

##### Parameters
**Gtotal**: Net solar irradiation (W/m^2)

**Tmin**: minimum temperature (degrees C)

**Tmax**: maximum temperature (degrees C)

**Tav**: average temperature (degrees C)

**Tday_length**: daytime length of the location (hours)

**T_ref**: room temperature is the reference temp (25 C)

**B, g c1, c2, c2**: parameters of solar panel technology

Read in raster files for model
```{r}
# 2020 projections
solar_stack_mask_20 <- stack("./generated_files/solar_stack_mask_20.tif") # Gtotal
tmin_stack_mask_20 <- stack("./generated_files/tmin_stack_mask_20.tif") # tmin
tmax_stack_mask_20 <- stack("./generated_files/tmax_stack_mask_20.tif") # tmax
temp_stack_mask_20 <- stack("./generated_files/temp_stack_mask_20.tif") # tac
```

```{r, warning = FALSE}
# run solar energy output model
solar_energy <- energy_output(Gtotal = solar_stack_mask_20,
                             Tmin = tmin_stack_mask_20,
                             Tmax = tmax_stack_mask_20,
                             Tav = temp_stack_mask_20,
                             Tday_length = 12,
                             g = 0.05*10^-3,
                             B = -4.5*10^-3)

# rename bands
names(solar_energy) <- c("jan", "feb", "mar", "apr", "may", "jun", 
                    "jul", "aug", "sep", "oct", "nov", "dec")
```

##### Projected Solar Panel Energy Output in CA for 2020:
```{r}
plot(solar_energy)
```

### Visualization
```{r}
# average across the raster stack to get annual average
solar_energy_av <- mean(solar_energy, na.rm = TRUE)
plot(solar_energy_av)

# extract average energy output data for each county and join to counties shapefile
county_energy <- raster::extract(solar_energy_av, ca_counties, fun=mean)

# give ca counties shapefile and county_energy matching row ids and join
ca_counties_id <- ca_counties %>% 
  mutate(num_id = row_number()) 

#tidy
county_energy_sp <- as.data.frame(county_energy) %>% 
  mutate(num_id = row_number()) %>% 
  left_join(ca_counties_id, by = "num_id") %>% 
  mutate(county = str_to_title(str_remove(ID, "california,"))) %>% 
  rename(energy_output = V1) %>% 
  dplyr::select(-num_id) %>% 
  arrange(energy_output) %>% 
  mutate(score = row_number())

# graph
ggplot(st_as_sf(county_energy_sp)) +
  geom_sf(aes(fill = energy_output)) +
 scale_fill_paletteer_c("ggthemes::Temperature Diverging", 
                        limits = c(32,44),
                        breaks = c(32, 34, 36, 38, 40, 42)) +
  labs(
    fill = "Energy Ouput (W/m^2)"
  ) +
  theme_classic() +
  scale_size(guide = "none")

```


## Sensitivity Analysis



## Projected Solar Land Suitability Model


## Economic Model