---
title: "Model Example"
author: "AnnaClaire Marley"
date: "6/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Rmarkdown runs through the solar suitability model with sample data cleaned and processed in the data_cleaning.Rmd

Load libraries
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(raster)
library(maps)
library(sf)
library(paletteer)
library(stringr)
library(RColorBrewer)
library(ggspatial)
library(here)
```

**Path to functions for the model**
```{r}
model_path <- "./Functions/model/"
```

# CRS
```{r}
# Assign projection: NAD83
NAD83 <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83
+units=m +no_defs +ellps=GRS80 +towgs84=0,0,0"
```

# CA counties

```{r}
## counties ##
ca_counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE)) %>% 
  filter(str_detect(ID, 'california'))

# make the polygon the same CRS as rasters
county_project <- st_transform(ca_counties, NAD83)
```

## Projected Solar Panel Energy Output Model

```{r}
# source function
source(paste0(model_path, "energy_output.R"))
```

##### Parameters
**Gtotal**: Net solar irradiation (W/m^2)

**Tmin**: minimum temperature (degrees C)

**Tmax**: maximum temperature (degrees C)

**Tav**: average temperature (degrees C)

**Tday_length**: daytime length of the location (hours)

**T_ref**: room temperature is the reference temp (25 C)

**B, g c1, c2, c2**: parameters of solar panel technology

Read in raster files for model
```{r}
# 2020 projections
solar_stack_mask_20 <- stack("./generated_files/solar_stack_mask_20.tif") # Gtotal
tmin_stack_mask_20 <- stack("./generated_files/tmin_stack_mask_20.tif") # tmin
tmax_stack_mask_20 <- stack("./generated_files/tmax_stack_mask_20.tif") # tmax
temp_stack_mask_20 <- stack("./generated_files/temp_stack_mask_20.tif") # tac
```

```{r, warning = FALSE}
# run solar energy output model
solar_energy <- energy_output(Gtotal = solar_stack_mask_20,
                             Tmin = tmin_stack_mask_20,
                             Tmax = tmax_stack_mask_20,
                             Tav = temp_stack_mask_20,
                             Tday_length = 12,
                             g = 0.05*10^-3,
                             B = -4.5*10^-3)

# rename bands
names(solar_energy) <- c("jan", "feb", "mar", "apr", "may", "jun", 
                    "jul", "aug", "sep", "oct", "nov", "dec")
```

##### Projected Solar Panel Energy Output in CA for 2020:
```{r}
plot(solar_energy)
```

### Visualization
```{r}
# average across the raster stack to get annual average
solar_energy_av <- mean(solar_energy, na.rm = TRUE)
plot(solar_energy_av)

# extract average energy output data for each county and join to counties shapefile
county_energy <- raster::extract(solar_energy_av, ca_counties, fun=mean)

# give ca counties shapefile and county_energy matching row ids and join
ca_counties_id <- ca_counties %>% 
  mutate(num_id = row_number()) 

#tidy
county_energy_sp <- as.data.frame(county_energy) %>% 
  mutate(num_id = row_number()) %>% 
  left_join(ca_counties_id, by = "num_id") %>% 
  mutate(county = str_to_title(str_remove(ID, "california,"))) %>% 
  rename(energy_output = V1) %>% 
  dplyr::select(-num_id) %>% 
  arrange(energy_output) %>% 
  mutate(score = row_number())

# graph
ggplot(st_as_sf(county_energy_sp)) +
  geom_sf(aes(fill = energy_output)) +
 scale_fill_paletteer_c("ggthemes::Temperature Diverging", 
                        limits = c(32,44),
                        breaks = c(32, 34, 36, 38, 40, 42)) +
  labs(
    fill = "Energy Ouput (W/m^2)"
  ) +
  theme_classic() +
  scale_size(guide = "none")

```


## Sensitivity Analysis

## Projected Solar Land Suitability Model

### Source functions
```{r}
source(paste0(model_path,"suitable_land_CA.R"))
source(paste0(model_path,"suitable_land_counties.R"))
```

### Run functions with 2020 data
```{r}
# Read 2020 land use/land cover raster
land_cover_2020 <- raster("generated_files/land_cover_2020.tif")

# Land suitability raster for all CA 
suitable_2020 <- suitable_land_CA(land_cover_raster = land_cover_2020)

# Use raster for all CA to get land suitability by county
county_maps_2020 <- suitable_land_counties(suitable_raster = suitable_2020, year = 2020)
county_maps_2020
```

### Run functions with 2050 data
```{r}
# Read 2050 land use/land cover raster
land_cover_2050 <- raster("generated_files/land_cover_2050.tif")

# Land suitability raster for all CA 
suitable_2050 <- suitable_land_CA(land_cover_raster = land_cover_2050)

# Use raster for all CA to get land suitability by county
county_maps_2050 <- suitable_land_counties(suitable_raster = suitable_2050, year = 2050)
county_maps_2050
```


## Economic Model

### Read-in data
```{r}
#Read in property tax shape file
raw_df <- read_sf(dsn = "generated_files", layer = "property_tax")

```

### Calculate unit land cost
```{r}
land_value_2019 <- raw_df 

# make the CRS consistent
land_value_2019 <- st_transform(land_value_2019, NAD83) %>% 
  rename(county = strCounty)
```

```{r}

#ploting the baseline map
ggplot(land_value_2019) +
  geom_sf(aes(fill = unitLand), color = NA) +
  scale_fill_paletteer_c("ggthemes::Orange") +
  scale_size(guide = "none") + #Removing size legend
  labs(fill = "Land Value ($/m^2)")

ggsave("land_value.png", path = "./Images")
  
```

### Calculate levelized cost of electricity (LCOE)
```{r}
#creating a data.frame storing the average solar output

solar_energy_simple <- county_energy_sp %>% 
  dplyr::select(-geom) %>% 
  mutate(strCounty = str_to_title(str_remove(ID, "california,"))) %>% 
  dplyr::select(-ID) %>% 
  rename(score_gen = score)


df_LCOE <- full_join(solar_energy_simple, land_value_2019)
```

```{r}
source(paste0(model_path,"LCOE.R"))

df_LCOE <- df_LCOE %>% 
  mutate(LCOE = LCOE(gen = energy_output, size = 20000, total_om = 8.12, instant_cost = 1585, installed_cost = 1861, land_cost = unitLand)) %>% 
  st_as_sf() %>% 
  arrange(-LCOE) %>% 
  mutate(score_econ = row_number())

#final_LCOE <- mapply(LCOE, gen = df_LCOE$energy_output, MoreArgs = list(size = 20000, total_om = 8.12, #instant_cost = 1742, installed_cost = 1861))

write_csv(df_LCOE, "generated_files/LCOE.csv")
```


### Visualization
```{r}
ggplot(df_LCOE) +
  geom_sf(aes(fill = LCOE), size = 0.2) +
  scale_fill_paletteer_c("scico::hawaii", direction = -1) +
  scale_size(guide = "none") + #Removing size legend
  labs(fill = "LCOE ($/MWh)")

ggsave("LCOE.png", path = "./Images")
  
```

## Final scores

### Source function
```{r}
source(paste0(model_path, "final_score.R"))
```

### Final score for year 2020
```{r}
# Get scores from 3 submodels
# Energy generation
score_gen_df <- as.data.frame(county_energy_sp) %>% 
  dplyr::select(county, score) %>% 
  rename(County = county, score_gen = score)

# Economic



# Land suitability
score_land_df <- county_maps_2020[[4]]
  
# Use score outputs from the 3 submodels as inputs for the final score function
final_score_df <-
  final_score(score_gen = score_gen_df,
              score_econ = score_econ_df,
              score_land = score_land_df)

# Look at map
final_score_df[[2]]
```



