---
title: "Solar Insolation"
author: "AnnaClaire Marley"
date: "5/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
# load packages
library(raster)
library(maps)
library(sf)
library(tidyverse)
library(paletteer)
library(stringr)
library(RColorBrewer)
```

## Calculate projected solar energy output in CA given solar radiation projections

#### **Workflow**

**Clean and process**

* Read in solar insolation raster data projected for California in 2050 
* Reproject, crop, and mask to California
* Read in average, max, and min daily temperature data projected for California in 2050
* Reproject, crop, and mask to California

**Run solar energy output model**

```{r}
# Create a Root to Extract Files 
if(Sys.info()["sysname"] == "Windows"){ #If operating system is "Windows" use G: Drive
  team_path <- "G:/"
} else { # If it is not (i.e. Mac), use the Google Drive File Stream
  team_path <- "/Volumes/GoogleDrive/"
}

#Create the Full File Path
path <- paste0(team_path, "Shared drives/ESM232/Data/in")

```

**CRS**
```{r}
# Assign projection: NAD83
NAD83 <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83
+units=m +no_defs +ellps=GRS80 +towgs84=0,0,0 "
```

## Clean and Process Data

```{r}

# function to read in raster data, create a raster stack, and rename files to months of the year
# folder_path: string of name of the folder as "/folder_name"

process_rasters = function(folder_path){
  
  #generate a list of input rasters 
  #pattern = "*.tif$" - filters for main raster files only and skips any associated files (e.g.   world files)
  rasters <- list.files(paste0(path, folder_path) , 
                             pattern = "*.tif$")
  
  #create a raster stack from the input raster files 
  raster_stack <- raster::stack(paste0(paste0(path, paste0(folder_path, "/")), rasters))
  
  # rename bands
  names(raster_stack) <- c("jan", "feb", "mar", "apr", "may", "jun", 
                        "jul", "aug", "sep", "oct", "nov", "dec")
  
  return(raster_stack)
}

```

**Make California shapefile**

```{r}
## State border##
# grab US states from maps package
california <- st_as_sf(maps::map("state", plot = FALSE, fill = TRUE)) %>% 
  filter(ID == "california")

# make the polygon the same CRS as rasters
california_project <- st_transform(california, NAD83)

## counties ##
ca_counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE)) %>% 
  filter(str_detect(ID, 'california'))

# make the polygon the same CRS as rasters
county_project <- st_transform(ca_counties, NAD83)

```

**Crop and mask rasters to California **

```{r}
# function to reproject, crop and mask all rasters to california

# raster: the raster files
# project: the projection you want to reproject the raster to
# polygon: california_project
ca_crop_mask = function(raster, project, polygon){
  
  reproject <- projectRaster(raster, crs = project)
  crop <- crop(reproject, polygon)
  mask <- mask(crop, polygon)
  
  return(mask)
}
```

#### Net surface radiation (W/m^2)

* Model: CCSM3
* Emission scenario: A2
* Year: 2050


```{r}
# net average surface radiation 
solar_stack <- process_rasters(folder_path = "/solar_2050")

# crop and mask to ca
solar_stack_mask <- ca_crop_mask(solar_stack, NAD83, california_project)

# plot
plot(solar_stack_mask)

```

### Temperature (C)

* Model: CCSM3
* Emission scenario: A2



#### Average Daily 

```{r}
#generate a list of input rasters 
temp_stack <- process_rasters(folder_path = "/tav_2050")

# crop and mask to ca
temp_stack_mask <- ca_crop_mask(temp_stack, NAD83, california_project)

# plot
plot(temp_stack_mask)
```

#### Average Maximum 

```{r}
# max temp
tmax_stack <- process_rasters(folder_path = "/tmax_2050")

# crop and mask to ca
tmax_stack_mask <- ca_crop_mask(tmax_stack, NAD83, california_project)

# plot
plot(tmax_stack_mask)
```

#### Average Minimum 

```{r}
# max temp
tmin_stack <- process_rasters(folder_path = "/tmin_2050")

# crop and mask to ca
tmin_stack_mask <- ca_crop_mask(tmin_stack, NAD83, california_project)

# plot
plot(tmin_stack_mask)
```


### Projected Energy Output Model

##### Parameters
**Gtotal**: Net solar irradiation (W/m^2)

**Tav**: average temperature (K)

**DTR**: daily temperature range - average temperature difference of the month

**Gday**: day-time insolation - Gtotal is total solar insolation (day + night) but we only need day time 

**Tday_length**: daytime length of the location (hours)

**e_cell/e_ref**: efficiency of the photo voltaic (PV) panel

**T_ref**: room temperature is the reference temp (25 C)

**T_cell**: temperature of the solar panel

**B, g c1, c2, c2**: parameters 

##### Equations

DTR = Tmax - Tmin

Gday = Gtotal * (24/Tday_length)

Tday = T + DIR/4

Tcell = c1 + c2(Tday) + c3(Gday)

eff = 1 - B(T_cell - T_ref) + g(log(Gday))

Ppv = Gday(e_cell)


Energy output equation
```{r}

# function to calculate energy output for each month
energy_output = function(Gtotal, Tmin, Tmax, 
                         Tav, Tday_length,  
                         B, g, Tref = 25, c1 = -3.75, 
                         c2 = 1.14, c3 = 0.0175){
  
  # calculate the average daily temperature range 
  DTR = Tmax - Tmin
  
  # Calculate average daytime temperature
  Tday = Tav + DTR/4
  
  # calculate daytime solar radiation
  Gday = Gtotal*(24/Tday_length)
  
  # calculate temperature of PV cell 
  Tcell = c1 + c2*Tday + c3*Gday
  
  # calculate efficiency of solar cells
  eff = 1 - B*(Tcell - Tref) + g*log10(Gday)
  
  # calculate solar output
  solar_output = Gday*eff*0.2
  
  return(solar_output)
  
}

```

Run solar energy output model

```{r}

solar_energy <- energy_output(Gtotal = solar_stack_mask,
                             Tmin = tmin_stack_mask,
                             Tmax = tmax_stack_mask,
                             Tav = temp_stack_mask,
                             Tday_length = 12,
                             g = 0.05*10^-3,
                             B = -4.5*10^-3)

# rename bands
names(solar_energy) <- c("jan", "feb", "mar", "apr", "may", "jun", 
                    "jul", "aug", "sep", "oct", "nov", "dec")
plot(solar_energy)


```

Collapse to year average
```{r}
# average across the raster stack 
solar_energy_av <- mean(solar_energy, na.rm = TRUE)
plot(solar_energy_av)

# extract average energy output data for each county and join to counties shapefile
county_energy <- raster::extract(solar_energy_av, ca_counties, fun=mean)

# give ca counties shapefile and county_energy matching row ids and join
ca_counties_id <- ca_counties %>% 
  mutate(num_id = row_number()) 

#tidy
county_energy_sp <- as.data.frame(county_energy) %>% 
  mutate(num_id = row_number()) %>% 
  left_join(ca_counties_id, by = "num_id") %>% 
  mutate(county = str_to_title(str_remove(ID, "california,"))) %>% 
  rename(energy_output = V1) %>% 
  select(-num_id) %>% 
  arrange(energy_output) %>% 
  mutate(score = row_number())

# graph

ggplot(st_as_sf(county_energy_sp)) +
  geom_sf(aes(fill = energy_output)) +
 scale_fill_paletteer_c("ggthemes::Temperature Diverging", 
                        limits = c(32,44),
                        breaks = c(32, 34, 36, 38, 40, 42)) +
  labs(
    fill = "Energy Ouput (W/m^2)"
  ) +
  theme_classic() +
  scale_size(guide = "none")

```

## Visualizations

% change is solar insolation
```{r}
s_2020 <- raster("./generated_files/solar_raster_2020.tif")
s_2050 <- raster("./generated_files/solar_raster_2050.tif")

# calculate

per_change <- ((s_2050 - s_2020)/s_2020)*100

# convert to df for plotting 
per_change_df <- raster::as.data.frame(per_change, xy = TRUE)

cols <- brewer.pal(n = 10, "Spectral")

plot(per_change, col = rev(cols))

```


