---
title: "solar_radiation_SA"
author: "Yingfei Jiang"
date: "5/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
# load packages
library(raster)
library(maps)
library(sf)
library(tidyverse)
library(paletteer)
library(sensitivity)
library(pse)
```

**Path to functions for the model**
```{r}
model_path <- "./Functions/model/"
```

# CRS
```{r}
# Assign projection: NAD83
NAD83 <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83
+units=m +no_defs +ellps=GRS80 +towgs84=0,0,0"
```

# CA counties

```{r}
## counties ##
ca_counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE)) %>% 
  filter(str_detect(ID, 'california'))

# make the polygon the same CRS as rasters
county_project <- st_transform(ca_counties, NAD83)
```

## Projected Solar Panel Energy Output Model

```{r}
# source function
source(paste0(model_path, "energy_output.R"))
```

##### Parameters
**Gtotal**: Net solar irradiation (W/m^2)

**Tmin**: minimum temperature (degrees C)

**Tmax**: maximum temperature (degrees C)

**Tav**: average temperature (degrees C)

**Tday_length**: daytime length of the location (hours)

**T_ref**: room temperature is the reference temp (25 C)

**B, g c1, c2, c2**: parameters of solar panel technology

Read in raster files for model
```{r}
# 2020 projections
solar_stack_mask_20 <- stack("./generated_files/solar_stack_mask_20.tif") # Gtotal
tmin_stack_mask_20 <- stack("./generated_files/tmin_stack_mask_20.tif") # tmin
tmax_stack_mask_20 <- stack("./generated_files/tmax_stack_mask_20.tif") # tmax
temp_stack_mask_20 <- stack("./generated_files/temp_stack_mask_20.tif") # tac
```

```{r}
#Write a wrapper function for SA

max_solar_output = function(Gtotal, Tmin, Tmax, Tav, Tday_length, B, g, c1, c2, c3){
  solar_energy <- energy_output(Gtotal = Gtotal,
                             Tmin = Tmin,
                             Tmax = Tmax,
                             Tav = Tav,
                             Tday_length = Tday_length,
                             g = g,
                             B = B,
                             c1 = c1,
                             c2 = c2,
                             c3 = c3)
  solar_energy_2050 <- mean(solar_energy, na.rm = TRUE)
  county_energy <- raster::extract(solar_energy_2050, ca_counties, fun=mean)
  ca_counties_id <- ca_counties %>% 
    mutate(num_id = row_number()) 
  county_energy_sp <- as.data.frame(county_energy) %>% 
    mutate(num_id = row_number()) %>%
    left_join(ca_counties_id, by = "num_id") %>%
    rename(energy_output = V1)
  
  return(max(county_energy_sp$energy_output))

}
```

#LHS

```{r}
#Evaluating two material parameters
factors = c("B", "g", "c1", "c2", "c3")

#Run 2000 sets
nsets = 2000

#Choose distributions
q = rep("qnorm", 5)
q.arg = list(list(mean=-4.5*10^-3, sd=4.5*10^-3/4), list(mean=0.05*10^-3, sd=0.05*10^-3/4), list(mean=-3.75, sd=3.75/4), list(mean=1.14, sd=1.14/4), list(mean=0.0175, sd=0.0175/4))

#Generate samples from LHS
sens_materials_lhs = LHS(NULL, factors, nsets, q, q.arg)
sens_pars_lhs = get.data(sens_materials_lhs)


#Calculate and store results
tmp = mapply(FUN = max_solar_output, B = sens_pars_lhs$B, g = sens_pars_lhs$g, c1 = sens_pars_lhs$c1, c2 = sens_pars_lhs$c2, c3 = sens_pars_lhs$c3, MoreArgs = list(Gtotal = solar_stack_mask, Tmin = tmin_stack_mask, Tmax = tmax_stack_mask, Tav = temp_stack_mask, Tday_length = 12))

sens_result_lhs = matrix(tmp)
colnames(sens_result_lhs) = c("maximum_output")


```

```{r}
sens_materials_lhs = pse::tell(sens_materials_lhs, t(sens_result_lhs), res.names = c("maximum_output"))
# x-y parameter plot
pse::plotscatter(sens_materials_lhs, col="blue", cex=5)
# cummulative distribution
pse::plotecdf(sens_materials_lhs,  col="blue", cex=5)
# partial rank correlation
pse::plotprcc(sens_materials_lhs, ylab = "Correlation")
sens_materials_lhs$prcc
```

```{r}
sens_result_lhs = as.data.frame(sens_result_lhs) %>% 
  gather(value="value", key="maximum_output")

ggplot(sens_result_lhs, aes(y = maximum_output))+
  geom_boxplot()+
  labs(y="Maximum Energy Output (W/m^2)")+
  theme_minimal()
```