---
title: "land_cover"
author: "Anna Calle"
date: "5/17/2020"
output: html_document
---

Data saved in google drive through file stream is separated into:
- Raw: raw data as downloaded
- In: renamed data used in R
- Out: data saved from R

Data required to run functions has been added to the generated_files function

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load required libraries 
library(tidyverse)
library(raster)
library(sf)
library(sp)
library(foreign)
library(maps)
library(mapdata)
library(ggplot2)
library(ggspatial)
library(ggrepel)
library(paletteer)
``` 

Set-Up for Downloading Data Using Google Drive File Stream
```{r}
# Create a Root to Extract Files 
if(Sys.info()["sysname"] == "Windows"){ # If operating system is "Windows" use G: Drive
  team_path <- "G:/"
} else { # If it is not (i.e. Mac), use the Google Drive File Stream
  team_path <- "/Volumes/GoogleDrive/"
}

# Create the Full File Paths for Inputs and Outputs
In <- paste0(team_path, "Shared drives/ESM232/Data/in/") # cannot use "in" as name
out <- paste0(team_path, "Shared drives/ESM232/Data/out/")

```

CRS
```{r}
# Assign projection: NAD83
NAD83 <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83
+units=m +no_defs +ellps=GRS80 +towgs84=0,0,0"
```

# CA
```{r}
# grab US states from maps package
california <- st_as_sf(maps::map("state", plot = FALSE, fill = TRUE)) %>% 
  filter(ID == "california")

# make the polygon the same CRS as rasters
CA_proj <- st_transform(california, NAD83)

# Save shapefile (to read inside function)
st_write(CA_proj, paste0(out, "/CA_proj.shp"), delete_layer = TRUE)

```

# CA Counties
```{r}
# grab US counties from maps package
counties_CA <- st_as_sf(maps::map("county", "california", plot = FALSE, fill = TRUE)) %>% 
  mutate(County = str_to_title(str_remove(ID, "california,"))) %>% 
  dplyr::select(-ID)

# reproject, remove state from county names, and calculate area
counties_proj <- st_transform(counties_CA, NAD83) %>% 
  mutate(area = st_area(geom))

# Save shapefile
st_write(counties_proj, paste0(out, "/counties_proj.shp"), delete_layer = TRUE)
```

# Slope (no need to run)
```{r}
# Read files for CA
slope_a <- raster(paste0(In, "slope_a.tif"))
slope_b <- raster(paste0(In, "/slope_b.tif"))
slope_c <- raster(paste0(In, "/slope_c.tif"))
slope_d <- raster(paste0(In,"/slope_d.tif"))

# Merge files
slope <- merge(slope_a, slope_b, slope_c, slope_d, progress = 'text')

# Crop
slope_cropped <- crop(slope, CA_proj)

# Mask
slope_masked <-mask(slope_cropped, CA_proj)

# Change resolution
slope_small <- raster::resample(slope_masked, land_cover, method='bilinear')

# Plot
plot(slope_small)

# Save
writeRaster(x = slope_small, filename = paste0(out,"slope_CA.tif"), overwrite = T)
```




# Test functions with 2020 data
```{r}
source("suitable_land_CA.R")
source("suitable_land_counties.R")
# Read 2020 land use/land cover raster
land_cover_2020 <- raster("generated_files/land_cover_2020.tif")

# Test land suitability for all CA function
suitable_2020 <- suitable_land_CA(land_cover_raster = land_cover_2020)

# Test mapping by county function
county_maps_2020 <- suitable_land_counties(suitable_raster = suitable_2020, year = 2020)
county_maps_2020
```

# Test functions with 2050 data
```{r}
# Read 2050 land use/land cover raster
land_cover_2050 <- raster("generated_files/land_cover_2050.tif")

# Test land suitability for all CA function
suitable_2050 <- suitability_fun(land_cover_raster = land_cover_2050)

# Test mapping by county function
county_maps_2050 <- map_suitability_fun(suitable_raster = suitable_2050, year = 2050)
county_maps_2050
```

```{r}
# Adding ranks to LCOE files
LCOE <- read_csv("generated_files/LCOE.csv") 

# Remove geometry and unnecessary columns
all_scores <- as.data.frame(LCOE) %>% 
  dplyr::select(-geometry, -strCounty, -energy_output, -unitLand, -LCOE) %>% 
  rename(score_land = rank_land)

# Separate gen scores
score_gen <- all_scores %>% 
  dplyr::select(County, score_gen)

# Separate econ scores
score_econ <- all_scores %>% 
  dplyr::select(County, score_econ)
```

# Final score function
```{r}
source("final_score.R")
# Test function
final_score_df <-
  final_score(score_gen, score_econ, county_maps_2020[[4]])

# Look at map
final_score_df[[2]]


```

